% Finding Groups within data Module 2 For Coursera Strategic Analytics
% Cesar Torres
% Created: 01/28/2016
% Updated: 01/28/2016 14:00

#Example 1

Let's have a look to the our variables and see some summary statistics.
ADS: Average Daily Sales
CV: Coefficient of Variation

```{r ShowData, echo = FALSE}
#The Gather2 has loaded the data into variable data
str(data)
summary(data)

```

Finding Groups by Visual Inspection
-----------------------------------

let's plot our data to see if we can indetify groups visually

```{r simplePlot, echo=FALSE}
plot(data$CV, data$ADS, main = "SKU Example", ylab = "Average Daily Sales", xlab = "Coefficient of Variation")
#draw a vertical line to separate the first group we detected visually
abline(v = 0.2, col = "red")
#draw an horizontal line to sepaate the second group we detected visually
abline(h = 4, col = "red")
#label the groups we detected visually
text(0.15, 9.7, "Horses", col = "red")
text(0.65, 9, "Wild Bulls", col = "red")
text(0.8, 2, "Crickets", col = "red")

```

Finding Groups using Hierarchical Clustering
--------------------------------------------
Now let's try to find groups using hierarchical clustering and check is we obtain similar results. We select to create three clusters to keep it simple for the business.

```{r hierarchicalClustering, echo = FALSE}
#Create copy of data 
testdata <- data
#In order to compare the data we need to normalize the data using the scale function
testdata <- scale(testdata)
#Compute the distances of all observations in our dataset using the euclidean method.
d <- dist(testdata, method = "euclidean")
#We use the hclust function by passing the distances between the data points using the ward.D method.
hcward <- hclust(d, method = "ward.D")
#We waqnt to pass the number of Cluster we want to select. It is important to limit this number to groups that business can understand.
data$groups <- cutree(hcward, k = 3)
```

Visualizing the Groups with more advance method
-----------------------------------------------

Now let's try to do a better visualizatio of the groups we found.

```{r advanceGroupVisual, echo = FALSE, warning = FALSE}
#Install the lattice package if not already installed.
install.packages("lattice")
#load the lattice package
library(lattice)
xyplot(ADS ~ CV, main = "After Clustering", type = "p", group=groups, data=data, #define the groups to be diffentiated
        auto.key = list(title = "Group", space = "left", cex = 1.0, just = 0.95),#produce the legend
        par.settings = list(superpose.line = list(pch = 0:18, cex = 1)),#pass the display settigns
		col=c('blue', 'green', 'red'))#colors for plotting the points

```