% Finding Groups within data Module 2 For Coursera Strategic Analytics
% Cesar Torres
% Created: 01/28/2016
% Updated: 01/28/2016 14:00

#Example 1

```{r getdata, echo = FALSE}
#Gather dataset for Finding Groups Module 2 Example 1
source("./StrategicAnalytics/Gather2.R")
```


Let's have a look to the our variables and see some summary statistics.
ADS: Average Daily Sales
CV: Coefficient of Variation

```{r ShowData, echo = FALSE}
#The Gather2 has loaded the data into variable data
str(data)
summary(data)

```



Finding Groups by Visual Inspection
-----------------------------------

let's plot our data to see if we can indetify groups visually

```{r simplePlot, echo=FALSE}
plot(data$CV, data$ADS, main = "SKU Example", ylab = "Average Daily Sales", xlab = "Coefficient of Variation")
#draw a vertical line to separate the first group we detected visually
abline(v = 0.2, col = "red")
#draw an horizontal line to sepaate the second group we detected visually
abline(h = 4, col = "red")
#label the groups we detected visually
text(0.15, 9.7, "Horses", col = "red")
text(0.65, 9, "Wild Bulls", col = "red")
text(0.8, 2, "Crickets", col = "red")

```

Finding Groups using Hierarchical Clustering
--------------------------------------------
Now let's try to find groups using hierarchical clustering and check is we obtain similar results. We select to create three clusters to keep it simple for the business.

```{r hierarchicalClustering, echo = FALSE}
#Create copy of data 
testdata <- data
#In order to compare the data we need to normalize the data using the scale function
testdata <- scale(testdata)
#Compute the distances of all observations in our dataset using the euclidean method.
d <- dist(testdata, method = "euclidean")
#We use the hclust function by passing the distances between the data points using the ward.D method.
hcward <- hclust(d, method = "ward.D")
#We waqnt to pass the number of Cluster we want to select. It is important to limit this number to groups that business can understand.
data$groups <- cutree(hcward, k = 3)
```

Visualizing the Groups with more advance method
-----------------------------------------------

Now let's try to do a better visualizatio of the groups we found.

```{r advanceGroupVisual, echo = FALSE}
#Install the lattice package if not already installed.
install.packages("lattice")
#load the lattice package
library(lattice)
xyplot(ADS ~ CV, main = "After Clustering", type = "p", group=groups, data=data, #define the groups to be diffentiated
        auto.key = list(title = "Group", space = "left", cex = 1.0, just = 0.95),#produce the legend
        par.settings = list(superpose.line = list(pch = 0:18, cex = 1)),#pass the display settigns
		col=c('blue', 'green', 'red'))#colors for plotting the points

```

#Example 2 - Human Resources

```{r getdataExample2, echo = FALSE}
#Gather dataset for Finding Groups Module 2 Example 2
source("./StrategicAnalytics/Gather3.R")
```

Let's have a look to the our variables and see some summary statistics.
S: Satisfaction variable under scale of 0 to 1
LPE: Last project Evaluation by a client under scale of 0 to 1
NP: Number of Project worked by an employee in the last 12 months
ANH: Average number of hours worked on the last 12 months.
TIC: Time Expend in the Company by teh employee
Newborn: Take a number of 1 if the employee has a newborn within the last 12 months and 0 otherwise.

```{r ShowDataExample2, echo = FALSE}
#The Gather2 has loaded the data into variable data
str(data)
summary(data)

```

Finding Groups using Hierarchical Clustering Example 2
------------------------------------------------------
Now let's try to find groups using hierarchical clustering and check is we obtain similar results. We select to create three clusters to keep it simple for the business. 
First we need to standarize the data by extracting the mean and divide by the standard deviation of the variable.

```{r hierarchicalClusteringExample2, echo = FALSE}
#Create copy of data 
testdata <- data
#In order to compare the data we need to normalize the data using the scale function
testdata <- scale(testdata)
#Compute the distances of all observations in our dataset using the euclidean method.
d <- dist(testdata, method = "euclidean")
#We use the hclust function by passing the distances between the data points using the ward.D method.
hcward <- hclust(d, method = "ward.D")
#We waqnt to pass the number of Cluster we want to select. It is important to limit this number to groups that business can understand.
data$groups <- cutree(hcward, k = 4) #Assign our points to to pur k=4 clusters
aggdata = aggregate(. ~ groups, data = data, FUN = mean) # The aggregate() function presents a summary of a statistic, broken down by one or more groups. Here we compute the mean of each variable for each group. 
```
